---
output:
  pdf_document: 
    highlight: kate
fontsize: 11pt
---

```{r setup, include=FALSE}
def.chunk.hook  <- knitr::knit_hooks$get("chunk")
knitr::knit_hooks$set(chunk = function(x, options) {
  x <- def.chunk.hook(x, options)
  ifelse(options$size != "normalsize", paste0("\n \\", options$size,"\n\n", x, "\n\n \\normalsize"), x)
})

knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, size = 'small')


# libraries
library(readxl)
library(broom)
library(plyr)
library(tidyverse)
library(ggplot2)
library(forcats)
library(tibble)
library(dplyr)
source('lib.R')
```

# Extension Experiments

In this markdown document we'll look at the extensions to the 'baseline'
transfer and persistence experiments.

## Camera settings data

In this experiment different camera settings were compared. The question 
being whether different camera settings, labeled 'C1' and 'C2', have an
influence on the particle counts.


```{r readEm}
# Read in Em's data
my.dat = read_xlsx("dat/StrEmCleanDataset.xlsx", sheet = "DataEm", trim_ws = TRUE, col_types = "text")

# Fix the column name for Time - seems to have extra spaces
my.dat = rename(my.dat, Mass = `Mass (g)`)
my.dat = rename(my.dat, TransferTime = `TransferTime (s)`)
my.dat = rename(my.dat, PersistenceTime = `PersistenceTime (min)`)
#colnames(my.dat)

# only use the Transfer data (PersistenceTime == "0") 
transfer.dat = my.dat %>% 
  filter(PersistenceTime == 0) %>% 
  select(c("Substrate", "ObservationType", "Count", "Mass", "TransferTime", "Experiment", "Replicate", "Note"))

# also remove expt 16 as it is incomplete
#transfer.dat = transfer.dat[as.numeric(transfer.dat$Experiment) < 16,]

knitr::kable(head(transfer.dat, 20), caption = "'Head' of EM's Transfer Data")

```



```{r mungeEM}
# fix column types
transfer.dat$Count = as.numeric(transfer.dat$Count)
transfer.dat$Experiment = as.character(transfer.dat$Experiment)
transfer.dat$Replicate = as.character(transfer.dat$Replicate)

# add a photoID column to keep track of which datapoint is which
photo.id = rep(paste0("P", seq(1:5)), as.integer(nrow(transfer.dat)/5+1))
transfer.dat = data.frame(PhotoID = photo.id[1:nrow(transfer.dat)], transfer.dat)

# split out each set of photos...
p1 = transfer.dat %>% filter(PhotoID == "P1") %>% select(Mass,TransferTime,Substrate,Experiment,Replicate,Note,Count)
p2 = transfer.dat %>% filter(PhotoID == "P2") %>% select(Mass,TransferTime,Substrate,Experiment,Replicate,Note,Count)
p3 = transfer.dat %>% filter(PhotoID == "P3") %>% select(Mass,TransferTime,Substrate,Experiment,Replicate,Note,Count)
p4 = transfer.dat %>% filter(PhotoID == "P4") %>% select(Mass,TransferTime,Substrate,Experiment,Replicate,Note,Count)
p5 = transfer.dat %>% filter(PhotoID == "P5") %>% select(Mass,TransferTime,Substrate,Experiment,Replicate,Note,Count)

# ...and merge
merge.dat = data.frame(p5[,c("Mass","TransferTime","Substrate","Experiment","Replicate", "Note")], p1$Count, p2$Count, p3$Count, p4$Count, p5$Count)
names(merge.dat) = c("Mass","TransferTime","Substrate","Experiment","Replicate", "Note","P1","P2","P3","P4","P5")

# The above has the side-effect of marking P1-P3 images with the
# Mass and TransferTime of P5. useful for turning into long format

# convert to long format
tlong.dat = gather(merge.dat, Photo, Count, P1:P5)

# rename substrates
tlong.dat[tlong.dat$Substrate == 'Wool01', 'Substrate'] <- 'Wool'
tlong.dat[tlong.dat$Substrate == 'Nylo01', 'Substrate'] <- 'Nylon'
knitr::kable(head(tlong.dat, 12), caption = "'Head' of EM data")
```

## Mass Data

Let's start with the 60s data over the various masses.

```{r ttdat}
# select 60s transfer time
t60.dat = tlong.dat %>% 
  filter(TransferTime == 60) %>%
  group_by(Mass, TransferTime, Experiment, Substrate, Note, Photo) %>%
  summarise(N = n(), mean = mean(Count), se = stdErr(Count))
knitr::kable(head(t60.dat), caption = "'Head' of Camera Data")

```

```{r cwplots}

t60wl.dat = t60.dat %>% filter(Substrate == "Wool", Photo %in% c('P3', 'P4', 'P5'))

ggplot(t60wl.dat, aes(x = numericFactorOrder(Mass), y = mean, colour = Note)) +
  labs(title = "Comparison of Camera Settings with Wool",
       subtitle = "Broken down by photo",
       x = "Mass (g)",
       y = "Particle Count",
       colour = "Camera\nSettings",
       caption = "Error bars: Std err") +
  geom_point(position = position_dodge(width = 0.3)) +
  geom_errorbar(aes(ymin = mean - se, ymax = mean + se), width = 0.08, position = position_dodge(width = 0.3)) +
  facet_grid(Photo ~ .) +
  facet_theme

```

```{r cnplots}

t60ny.dat = t60.dat %>% filter(Substrate == "Nylon", Photo %in% c('P3', 'P4', 'P5'))

ggplot(t60ny.dat, aes(x = numericFactorOrder(Mass), y = mean, colour = Note)) +
  labs(title = "Comparison of Camera Settings with Nylon",
       subtitle = "Broken down by photo",
       x = "Mass (g)",
       y = "Particle Count",
       colour = "Camera\nSettings",
       caption = "Error bars: Std err") +
  geom_point(position = position_dodge(width = 0.3)) +
  geom_errorbar(aes(ymin = mean - se, ymax = mean + se), width = 0.08, position = position_dodge(width = 0.3)) +
  facet_grid(Photo ~ .) +
  facet_theme

```

## Transfer Time

Compare camera settings across Transfer Time.

```{r timetrend}

m1000.dat = tlong.dat %>% 
  filter(Mass == 1000, TransferTime != 10) %>%
  group_by(Mass, TransferTime, Experiment, Substrate, Note, Photo) %>%
  summarise(N = n(), mean = mean(Count), se = stdErr(Count))

# remove single rep expts
m1000.dat = m1000.dat[m1000.dat$N > 1,]


knitr::kable(head(m1000.dat), caption = "'Head' of Camera Data")

```

```{r masswlplt}

m1000wl.dat = m1000.dat %>% filter(Substrate == "Wool", Photo %in% c('P3', 'P4', 'P5'))

ggplot(m1000wl.dat, aes(x = numericFactorOrder(TransferTime), y = mean, colour = Note)) +
  labs(title = "Comparison of Camera Settings with Wool",
       subtitle = "Broken down by photo",
       x = "Transfer Time (s)",
       y = "Particle Count",
       colour = "Camera\nSettings",
       caption = "Error bars: Std err") +
  geom_point(position = position_dodge(width = 0.3)) +
  geom_errorbar(aes(ymin = mean - se, ymax = mean + se), width = 0.08, position = position_dodge(width = 0.3)) +
  facet_grid(Photo ~ .) +
  facet_theme

```
```{r massnyplt}

m1000ny.dat = m1000.dat %>% filter(Substrate == "Nylon", Photo %in% c('P3', 'P4', 'P5'))

ggplot(m1000ny.dat, aes(x = numericFactorOrder(TransferTime), y = mean, colour = Note)) +
  labs(title = "Comparison of Camera Settings with Nylon",
       subtitle = "Broken down by photo",
       x = "Transfer Time (s)",
       y = "Particle Count",
       colour = "Camera\nSettings",
       caption = "Error bars: Std err") +
  geom_point(position = position_dodge(width = 0.3)) +
  geom_errorbar(aes(ymin = mean - se, ymax = mean + se), width = 0.08, position = position_dodge(width = 0.3)) +
  facet_grid(Photo ~ .) +
  facet_theme

```

```{r c1plt}
# compare camera settings by substrate type

c1.dat = t60.dat %>% filter(Note == 'C1', Photo %in% c('P3', 'P4', 'P5'))
  
ggplot(c1.dat, aes(x = numericFactorOrder(Mass), y = mean, colour = Substrate)) +
  labs(title = "Comparison of Camera Settings - C1",
       subtitle = "Broken down by photo and substrate",
       x = "Transfer Time (s)",
       y = "Particle Count",
       colour = "Camera\nSettings",
       caption = "Error bars: Std err") +
  geom_point(position = position_dodge(width = 0.3)) +
  geom_errorbar(aes(ymin = mean - se, ymax = mean + se), width = 0.08, position = position_dodge(width = 0.3)) +
  facet_grid(Photo ~ .) +
  facet_theme

```

```{r c2plt}
# compare camera settings by substrate type

c2.dat = t60.dat %>% filter(Note == 'C2', Photo %in% c('P3', 'P4', 'P5'))
  
ggplot(c2.dat, aes(x = numericFactorOrder(Mass), y = mean, colour = Substrate)) +
  labs(title = "Comparison of Camera Settings - C2",
       subtitle = "Broken down by photo and substrate",
       x = "Transfer Time (s)",
       y = "Particle Count",
       colour = "Camera\nSettings",
       caption = "Error bars: Std err") +
  geom_point(position = position_dodge(width = 0.3)) +
  geom_errorbar(aes(ymin = mean - se, ymax = mean + se), width = 0.08, position = position_dodge(width = 0.3)) +
  facet_grid(Photo ~ .) +
  facet_theme

```


## Do Camera Settings Make a Statistical Difference?

This is count data which won't be normally distributed, so we'll go
with the non-parametric Mann-Whitney test.

What we want to test is the effect of camera settings C1 & C2 on the 
particle counts on the substrates. We know that the counts are dependent
on the substrate and photo number at the very least and possible mass and
transfer time.

```{r stat1}
# get all data for photos 3, 4 & 5 exluding time 10
p345.dat = tlong.dat %>% 
  filter(TransferTime != 10, Photo %in% c('P3', 'P4', 'P5')) %>%
  select(Substrate, Photo, Mass, TransferTime, Note, Count)
# for all the combinations of substrate, mass, transfer time and photo
# calculate the wilcoxon p-value
wt.dat = p345.dat %>%
  group_by(Substrate, Mass, TransferTime, Photo) %>% 
  do(wt = wilcox.test(Count ~ Note, data = .)) %>% 
  tidy(wt) %>% 
  select(Substrate, Photo, Mass, TransferTime, statistic, p.value)

# as we done a lot of pairwise tests do a multiple hypothesis correction.
# Abitrarily chosen 'BH'. Need to check if that's apropriate here.
wt.dat = cbind(wt.dat, p.adj = p.adjust(wt.dat$p.value, method = 'BH'))
knitr::kable(head(wt.dat[order(wt.dat$p.adj),], 10), caption = "Most significantly affected counts")

```

What if we take out the transfer time and mass factors? From the transfer 
analysis we saw that there isn't really any relationship between time and
mass vs count.

```{r stat2}

# get all data for photos 3, 4 & 5 exluding time 10
p345.dat = tlong.dat %>% 
  filter(TransferTime != 10, Photo %in% c('P3', 'P4', 'P5')) %>%
  select(Substrate, Photo, Note, Count)
# for all the combinations of substrate and photo
# calculate the wilcoxon p-value
wt.dat = p345.dat %>% 
  group_by(Substrate, Photo) %>% 
  do(wt = wilcox.test(Count ~ Note, data = .)) %>% 
  tidy(wt) %>% 
  select(Substrate, Photo, statistic, p.value)

# as we done a lot of pairwise tests do a multiple hypothesis correction.
# Abitrarily chosen 'BH'. Need to check if that's apropriate here.
wt.dat = cbind(wt.dat, p.adj = p.adjust(wt.dat$p.value, method = 'BH'))
knitr::kable(head(wt.dat[order(wt.dat$p.adj),], 10), caption = "Most significantly affected counts")
```

Surprise! All the wool data are now significant even after correction. Let's see what that looks like.


```{r sigplot}

wsig.dat = p345.dat %>% filter(Substrate == 'Wool')

ggplot(wsig.dat, aes(x = Photo, y = Count, colour = Note)) +
  labs(title = "All Wool Particle Counts",
       subtitle = "by camera setting",
       x = "Image Number",
       y = "Particle Counts") +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(aes(colour = Note),position = position_jitterdodge(jitter.width = 0.1, dodge.width = 0.8), size = 0.9, shape = 21) +
  mytheme
```

So, does this mean that camera settings may need to be substrate-specific?
Maybe. More data required. Natch!

