---
title: "Persistence Experiments"
author: "Christian Cole (original author), Hervé Ménard (code editing), Niamh Nic Daéid (code editing)"
output: 
  pdf_document: 
    highlight: kate
fontsize: 11
---
```{r setup, include=FALSE}
def.chunk.hook  <- knitr::knit_hooks$get("chunk")
knitr::knit_hooks$set(chunk = function(x, options) {
  x <- def.chunk.hook(x, options)
  ifelse(options$size != "normalsize", paste0("\n \\", options$size,"\n\n", x, "\n\n \\normalsize"), x)
})

knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, size = 'small')
# libraries
library(readxl)
library(broom)
library(plyr)
library(tidyverse)
library(ggplot2)
library(ggpubr)
library(forcats)
library(tibble)
library(dplyr)
source('lib.R')
```

```{r themes, echo=FALSE}

mytheme = theme_minimal() +
  theme(panel.grid.minor.x = element_blank(),
        panel.grid.major.x = element_blank(),
        axis.title = element_text(size = 13, face = "bold"),
        axis.text = element_text(size = 11),
        legend.title = element_blank())

inset_theme = theme_minimal() +
  theme(plot.background = element_rect(fill = 'white', colour = 'white'),
        panel.grid.minor = element_blank(),
        panel.grid.major.x = element_blank(),
        axis.title = element_text(size = 8),
        axis.text = element_text(size = 8),
        legend.title = element_blank(),
        legend.position = 'NA')


```

# Analysis of Persistence Experiments

## Dataset A
```{r persA}
# Read in A data
my.dat = read_xlsx("dat/Counts_data.xlsx", sheet = "A")

# use the persistence experiments for Groups 1 & 2.
persist.dat = my.dat %>% filter(
      Substrate != 'Cott01' & 
      Experiment %in% c(1,2) & 
      ObservationType != 'Ndata') %>%
  select(c("Substrate", "ObservationType", "Count", "Mass (g)", "TransferTime (s)",
           "PersistenceTime (min)", "Experiment", "Replicate")) %>%
  dplyr::rename(Mass = `Mass (g)`, 
         TransferTime = `TransferTime (s)`,
         PersistenceTime = `PersistenceTime (min)`)

# fix column types
persist.dat$Count = as.numeric(persist.dat$Count)
persist.dat$PersistenceTime = as.numeric(persist.dat$PersistenceTime)
persist.dat$Experiment = as.character(persist.dat$Experiment)
persist.dat$Replicate = as.character(persist.dat$Replicate)
# some errors in the data, fix
persist.dat[persist.dat$Substrate == 'wool01', 'Substrate'] <- 'Wool'
persist.dat[persist.dat$Substrate == 'Wool01', 'Substrate'] <- 'Wool'
persist.dat[persist.dat$Substrate == 'Nylo01', 'Substrate'] <- 'Nylon'
persist.dat[persist.dat$Substrate == 'Wool' & persist.dat$Experiment == '2', 'Substrate'] <- 'Nylon'
# summarise the count data
summ.dat = summarySE(persist.dat, measurevar = "Count", groupvars = c("Substrate", "PersistenceTime",
                                                                      "Experiment"))
knitr::kable(head(summ.dat), caption = "Summary of A Persistence Data")
```



```{r mlplots,fig.height=3.9}


p1 = ggplot(summ.dat, aes(x=PersistenceTime, y = Count, group = Substrate, colour =
                            Substrate, shape=Substrate)) +
  geom_point(position = position_dodge(0.3)) +
  geom_errorbar(aes(ymin = Count-se, ymax = Count+se), width = 0.08) +
  xlim(c(-5,1500)) +
  scale_y_continuous(breaks = seq(0,60,10)) +
  scale_color_brewer(palette = 'Paired') +
  labs(x = "Persistence Time (min)",
       y = "Particle Count",
       caption = "Error bars: std error") +
  mytheme +
  theme_pubr()

p2 = ggplot(summ.dat, aes(x=PersistenceTime, y = Count, group = Substrate, colour =
                            Substrate, shape=Substrate)) +
  geom_point(position = position_dodge(0.3)) +
  geom_errorbar(aes(ymin = Count-se, ymax = Count+se), width = 0.08) +
  scale_color_brewer(palette = 'Paired') +
  labs(x = "Persistence Time (min)",
       y = "Particle Count") +
  theme_pubr() +
  theme(axis.title = element_text(size = 8),
        axis.text = element_text(size = 8),
        legend.title = element_blank(),
        legend.position = 'NA')

p1 + annotation_custom(ggplotGrob(p2), xmin = 700, xmax = 1600, 
                       ymin = 20, ymax = 65)
```

Figure S1: Particle count for Nylon and Wool as receiver materials

```{r fits, echo=TRUE}
# ideas taken from http://douglas-watson.github.io/post/2018-09_exponential_curve_fitting/
# and https://dataconomy.com/2017/08/nonlinear-least-square-nonlinear-regression-r/
set.seed(12345)
ny.fit = nls(Count ~ SSasymp(PersistenceTime, Countf, Count0, log_alpha), 
             data = summ.dat, 
             subset = Substrate == 'Nylon')
ny.fit.err = summ.dat$Count-predict(ny.fit)
ny.nlm_error <- sqrt(mean(ny.fit.err^2)) 
ny.fit
```

The residual error is: `r signif(ny.nlm_error)`



```{r fitplt,fig.height=3.9}
pts = seq(0,max(summ.dat$PersistenceTime),length=200)
gg.fit = data.frame(y=predict(ny.fit, data.frame(PersistenceTime = pts)), x= pts)

p1 = ggplot(summ.dat[summ.dat$Substrate == 'Nylon',], aes(x = PersistenceTime, y = Count)) +
  labs(
       x = "Persistence Time (min)",
       y = "Particle Count") +
  geom_point(colour = '#a6cee3') + 
  geom_errorbar(aes(ymin = Count-se, ymax = Count+se), width = 0.08, data =
                  summ.dat[summ.dat$Substrate == 'Nylon',], colour = '#a6cee3') +
  geom_line(aes(x = x, y = y), data = gg.fit, colour = '#a6cee3') +
  xlim(c(-5,1500)) +
  mytheme + 
  theme_pubr()



p2 = ggplot(summ.dat[summ.dat$Substrate == 'Nylon',], aes(x = PersistenceTime, y = Count)) +
  labs(x = "Persistence Time (min)",
       y = "Particle Count") +
  geom_point(colour = '#a6cee3') + 
  geom_errorbar(aes(ymin = Count-se, ymax = Count+se), width = 0.08, data =
                  summ.dat[summ.dat$Substrate == 'Nylon',], colour = '#a6cee3') +
  geom_line(aes(x = x, y = y), data = gg.fit, colour = '#a6cee3') +
  theme_pubr() +
  theme(axis.title = element_text(size = 8),
        axis.text = element_text(size = 8),
        legend.title = element_blank(),
        legend.position = 'NA')

nplt <- p1 + annotation_custom(ggplotGrob(p2), xmin = 700, xmax = 1600, 
                       ymin = 20, ymax = 65)
nplt
```

Figure S2: Curve fitting for Nylon as receiver material.

```{r woolfit}
wl.fit = nls(Count ~ SSasymp(PersistenceTime, Countf, Count0, log_alpha), 
             data = summ.dat, 
             subset = Substrate == 'Wool')
wl.fit.err = summ.dat$Count-predict(wl.fit)
wl.nlm_error <- sqrt(mean(wl.fit.err^2)) 
wl.fit
```
Residual error:  `r signif(wl.nlm_error)`

```{r woolfitplt,fig.height=3.9}
# generate more points for smooth regression line
pts = seq(0,max(summ.dat$PersistenceTime),length=200)
gg.fit = data.frame(y=predict(wl.fit, data.frame(PersistenceTime = pts)), x= pts)


p1 = ggplot(summ.dat[summ.dat$Substrate == 'Wool',], aes(x = PersistenceTime, y = Count)) +
  labs(
       x = "Persistence Time (min)",
       y = "Particle Count",
       caption = "Error bars: std error") +
  geom_point(colour = '#1f78b4') + 
  geom_errorbar(aes(ymin = Count-se, ymax = Count+se), width = 0.08,
                data = summ.dat[summ.dat$Substrate == 'Wool',], colour = '#1f78b4') +
  geom_line(aes(x = x, y = y), data = gg.fit, colour = '#1f78b4') +
  xlim(c(-5,1500)) +
  mytheme +
  theme_pubr()

p2 = ggplot(summ.dat[summ.dat$Substrate == 'Wool',], aes(x = PersistenceTime, y = Count)) +
  labs(
       x = "Persistence Time (min)",
       y = "Particle Count") +
  geom_point(colour = '#1f78b4') + 
  geom_errorbar(aes(ymin = Count-se, ymax = Count+se), width = 0.08,
                data = summ.dat[summ.dat$Substrate == 'Wool',], colour = '#1f78b4') +
  geom_line(aes(x = x, y = y), data = gg.fit, colour = '#1f78b4') +
  theme_pubr() +
  theme(axis.title = element_text(size = 8),
        axis.text = element_text(size = 8),
        legend.title = element_blank(),
        legend.position = 'NA')

wplt = p1 + annotation_custom(ggplotGrob(p2), xmin = 700, xmax = 1600, 
                       ymin = 10, ymax = 40)
wplt
```

Figure S3: Curve fitting for Wool as receiver material.

```{r combplot,fig.height=4.5}
arrplot = ggarrange(nplt + rremove('x.title'), wplt, labels = c('A', 'B'), nrow = 2)
#ggsave("persistence.png", arrplot, height = 4.5, units = 'in')
arrplot
```

Figure S4: Comparison between (A) Nylon and (B) Wool as receiver materials 

```{r}
summ.dat$Experiment <- as.numeric(summ.dat$PersistenceTime)
cmp.fit.dat = summ.dat %>% 
  group_by(Substrate) %>% 
  do(fit = nls(Count ~ SSasymp(PersistenceTime, Countf, Count0, log_alpha), data = .)) %>%
  ungroup %>%
  mutate(fitCoef = map(fit,tidy)) %>%
  select(Substrate, fitCoef) %>%
  unnest(fitCoef) %>%
  select(Substrate, term, estimate) %>% 
   spread(term, estimate) %>% 
   mutate(alpha = exp(log_alpha))
knitr::kable(cmp.fit.dat)
```

## Datasets B, C & D

No persistence data available. Data acquisition focussed on the transfer study only.

## Dataset E

```{r readEm}
# Read in E data
my.dat = read_xlsx("dat/Counts_data.xlsx", sheet = "E", trim_ws = TRUE, col_types = "text")


# use exp 7 & 8
persist.dat = my.dat %>% filter(
      Substrate != 'Cott01' & 
      Experiment %in% c(7,8) & 
      Replicate > 7 &
      ObservationType != 'Ndata') %>%
  select(c("Substrate", "ObservationType", "Count", "Mass (g)", 
           "PersistenceTime (min)", "Experiment", "Replicate", "Note")) %>%
  dplyr::rename(Mass = `Mass (g)`, 
         PersistenceTime = `PersistenceTime (min)`)

# fix column types
persist.dat$Count = as.numeric(persist.dat$Count)
persist.dat$PersistenceTime = as.numeric(persist.dat$PersistenceTime)
persist.dat$Experiment = as.character(persist.dat$Experiment)
persist.dat$Replicate = as.character(persist.dat$Replicate)

# summarise the count data
summ.dat = summarySE(persist.dat, measurevar = "Count",
                     groupvars = c("Substrate", "PersistenceTime", "Experiment", "Note"))
knitr::kable(summ.dat, caption = "Summary of E Persistence Data")

```

Insufficient number of replicates for comprehensive analysis: only 2 replicates for 0, 30, 60 and 1440 minutes, all the other results are single-point.

```{r plotE,fig.height=3.9}

summ.dat = persist.dat %>% 
  group_by(Substrate, PersistenceTime, Note) %>%
  summarise(N = n(), Mean = mean(Count), Max = max(Count), Min = min(Count))

ggplot(summ.dat, aes(x=PersistenceTime, y = Mean, colour = Note)) +
  geom_point() +
  geom_linerange(aes(ymin = Min, ymax = Max)) +
  scale_color_brewer(palette = 'Paired') +
  xlim(c(-10, 1500)) +
  labs(title = "Nylon Persistence",
       subtitle = "By different camera settings",
       x = "Persistence Time (min)",
       y = "Particle Count",
       caption = "Error bars: min-max") +
  mytheme +
  theme_pubr()
  
```
Figure S5: Nylon as receiver material, comparison between camera settings.
