---
title: "Persistence Experiments"
author: "Christian Cole"
date: "10/03/2020"
output: 
  pdf_document: 
    highlight: kate
fontsize: 11
---

```{r setup, include=FALSE}
def.chunk.hook  <- knitr::knit_hooks$get("chunk")
knitr::knit_hooks$set(chunk = function(x, options) {
  x <- def.chunk.hook(x, options)
  ifelse(options$size != "normalsize", paste0("\n \\", options$size,"\n\n", x, "\n\n \\normalsize"), x)
})

knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, size = 'small')
# libraries
library(readxl)
library(broom)
library(plyr)
library(tidyverse)
library(ggplot2)
library(forcats)
library(tibble)
library(dplyr)
source('lib.R')
```

```{r themes, echo=FALSE}

mytheme = theme_minimal() +
  theme(panel.grid.minor.x = element_blank(),
        panel.grid.major.x = element_blank(),
        axis.title = element_text(size = 13, face = "bold"),
        axis.text = element_text(size = 11),
        legend.title = element_blank())

inset_theme = theme_minimal() +
  theme(plot.background = element_rect(fill = 'white', colour = 'white'),
        panel.grid.minor = element_blank(),
        panel.grid.major.x = element_blank(),
        axis.title = element_text(size = 8),
        axis.text = element_text(size = 8),
        legend.title = element_blank(),
        legend.position = 'NA')


```

# Analysis of Persistence Experiments

Go through experiment sets and extract the persistence data. The data are
defined by having "ObsType" != "Ndata".

## Expt 1

```{r persML}
# Read in ML's data
my.dat = read_xlsx("dat/20191030DatasetPaper.xlsx", sheet = "DataMl")

# only use the persistence experiments for Groups 1 & 2. Groups 3-6 do not 
# have enough replicates to be meaningful
persist.dat = my.dat %>% filter(
      Substrate != 'Cott01' & 
      Group %in% c(1,2) & 
      ObsType != 'Ndata') %>%
  select(c("Substrate", "ObsType", "Count", "Mass", "Duration", "Time", "Group", "Repeat")) %>%
  rename(ObservationType = ObsType, 
         TransferTime = Duration,
         PersistenceTime = Time,
         Experiment = Group,
         Replicate = Repeat)

# fix column types
persist.dat$Count = as.numeric(persist.dat$Count)
persist.dat$PersistenceTime = as.numeric(persist.dat$PersistenceTime)
persist.dat$Experiment = as.character(persist.dat$Experiment)
persist.dat$Replicate = as.character(persist.dat$Replicate)
# some errors in the data, fix
persist.dat[persist.dat$Substrate == 'wool01', 'Substrate'] <- 'Wool'
persist.dat[persist.dat$Substrate == 'Wool01', 'Substrate'] <- 'Wool'
persist.dat[persist.dat$Substrate == 'Nylo01', 'Substrate'] <- 'Nylon'
persist.dat[persist.dat$Substrate == 'Wool' & persist.dat$Experiment == '2', 'Substrate'] <- 'Nylon'

# summarise the count data
summ.dat = summarySE(persist.dat, measurevar = "Count", groupvars = c("Substrate", "PersistenceTime", "Experiment"))
knitr::kable(head(summ.dat), caption = "Summary of ML's Persistence Data")

```

The data covers a large period of time - potentially weeks - so the time
scale needs to be compressed somehow for ease of interpretation. Let's see
what we can do.

```{r mlplots}

p1 = ggplot(summ.dat, aes(x=PersistenceTime, y = Count, group = Substrate, colour = Substrate, shape=Substrate)) +
  geom_point(position = position_dodge(0.3)) +
  geom_errorbar(aes(ymin = Count-se, ymax = Count+se), width = 0.08) +
  xlim(c(-5,1500)) +
  scale_y_continuous(breaks = seq(0,60,10)) +
  labs(title = "Persistence Experiment",
       subtitle = "60s & 1000g Transfer",
       x = "Persistence Time (min)",
       y = "Particle Count",
       caption = "Error bars: std error") +
  mytheme

p2 = ggplot(summ.dat, aes(x=PersistenceTime, y = Count, group = Substrate, colour = Substrate, shape=Substrate)) +
  geom_point(position = position_dodge(0.3)) +
  geom_errorbar(aes(ymin = Count-se, ymax = Count+se), width = 0.08) +
  labs(x = "Persistence Time (min)",
       y = "Particle Count") +
  inset_theme

p1 + annotation_custom(ggplotGrob(p2), xmin = 700, xmax = 1600, 
                       ymin = 20, ymax = 65)
```

The data seems to be a smooth decay. Let's try some curve-fitting.

This function SSasymp() a self-starting fucntion which attempts to find
suitable initial parameters for perform a fit - here a non-linear least
squares.

Let's see the Nylon fit first

```{r fits, echo=TRUE}
# ideas taken from http://douglas-watson.github.io/post/2018-09_exponential_curve_fitting/
# and https://dataconomy.com/2017/08/nonlinear-least-square-nonlinear-regression-r/
set.seed(12345)
ny.fit = nls(Count ~ SSasymp(PersistenceTime, Countf, Count0, log_alpha), 
             data = summ.dat, 
             subset = Substrate == 'Nylon')
ny.fit.err = summ.dat$Count-predict(ny.fit)
ny.nlm_error <- sqrt(mean(ny.fit.err^2)) 
ny.fit
```

On first impressions the fit looks fine. The residual error is: `r signif(ny.nlm_error)`

A plot of the data and the fit will allow a visual comparison.

```{r fitplt}
pts = seq(0,max(summ.dat$PersistenceTime),length=200)
gg.fit = data.frame(y=predict(ny.fit, data.frame(PersistenceTime = pts)), x= pts)

p1 = ggplot(summ.dat[summ.dat$Substrate == 'Nylon',], aes(x = PersistenceTime, y = Count)) +
  labs(title = "Nylon Persistence",
       subtitle = "With non-linear least squares fit",
       x = "Persistence Time (min)",
       y = "Particle Count",
       caption = "Error bars: std error") +
  geom_point() + 
  geom_errorbar(aes(ymin = Count-se, ymax = Count+se), width = 0.08, data = summ.dat[summ.dat$Substrate == 'Nylon',]) +
  geom_line(aes(x = x, y = y), data = gg.fit) +
  xlim(c(-5,1500)) +
  mytheme

p2 = ggplot(summ.dat[summ.dat$Substrate == 'Nylon',], aes(x = PersistenceTime, y = Count)) +
  labs(x = "Persistence Time (min)",
       y = "Particle Count") +
  geom_point() + 
  geom_errorbar(aes(ymin = Count-se, ymax = Count+se), width = 0.08, data = summ.dat[summ.dat$Substrate == 'Nylon',]) +
  geom_line(aes(x = x, y = y), data = gg.fit) +
  inset_theme

p1 + annotation_custom(ggplotGrob(p2), xmin = 700, xmax = 1600, 
                       ymin = 20, ymax = 65)
```

The nylon data looks good, let's to the same with the wool data and compare.

```{r woolfit}
wl.fit = nls(Count ~ SSasymp(PersistenceTime, Countf, Count0, log_alpha), 
             data = summ.dat, 
             subset = Substrate == 'Wool')
wl.fit.err = summ.dat$Count-predict(wl.fit)
wl.nlm_error <- sqrt(mean(wl.fit.err^2)) 
wl.fit
```
Residual error:  `r signif(wl.nlm_error)`

```{r woolfitplt}
# generate more points for smooth regression line
pts = seq(0,max(summ.dat$PersistenceTime),length=200)
gg.fit = data.frame(y=predict(wl.fit, data.frame(PersistenceTime = pts)), x= pts)


p1 = ggplot(summ.dat[summ.dat$Substrate == 'Wool',], aes(x = PersistenceTime, y = Count)) +
  labs(title = "Wool Persistence",
       subtitle = "With non-linear least squares fit",
       x = "Persistence Time (min)",
       y = "Particle Count",
       caption = "Error bars: std error") +
  geom_point() + 
  geom_errorbar(aes(ymin = Count-se, ymax = Count+se), width = 0.08, data = summ.dat[summ.dat$Substrate == 'Wool',]) +
  geom_line(aes(x = x, y = y), data = gg.fit) +
  xlim(c(-5,1500)) +
  mytheme

p2 = ggplot(summ.dat[summ.dat$Substrate == 'Wool',], aes(x = PersistenceTime, y = Count)) +
  geom_point() + 
  geom_errorbar(aes(ymin = Count-se, ymax = Count+se), width = 0.08, data = summ.dat[summ.dat$Substrate == 'Wool',]) +
  geom_line(aes(x = x, y = y), data = gg.fit) +
  inset_theme

p1 + annotation_custom(ggplotGrob(p2), xmin = 700, xmax = 1600, 
                       ymin = 10, ymax = 40)
```

```{r}
cmp.fit.dat = summ.dat %>% 
  group_by(Substrate) %>% 
  do(fit = nls(Count ~ SSasymp(PersistenceTime, Countf, Count0, log_alpha), data = .)) %>% 
  tidy(fit) %>% 
  select(Substrate, term, estimate) %>% 
  spread(term, estimate) %>% 
  mutate(alpha = exp(log_alpha))
knitr::kable(cmp.fit.dat)
```

Interestingly the decay rate (alpha) is almost the same for the two materials.

## Expt 2, 3 & 4

No persistence data.

## Expt 5

The only other dataset with persistence results is EM. 

```{r readEm}
# Read in Em's data
my.dat = read_xlsx("dat/StrEmCleanDataset.xlsx", sheet = "DataEm", trim_ws = TRUE, col_types = "text")

# Fix the column name for Time - seems to have extra spaces
my.dat = rename(my.dat, Mass = `Mass (g)`)
my.dat = rename(my.dat, TransferTime = `TransferTime (s)`)
my.dat = rename(my.dat, PersistenceTime = `PersistenceTime (min)`)
#colnames(my.dat)

# only use the Transfer data (PersistenceTime == "0") 
# transfer.dat = my.dat %>% filter(PersistenceTime == 0) %>% select(c("Substrate", "ObservationType", "Count", "Mass", "TransferTime", "Experiment", "Replicate", "Note"))

# only use expts 7 & 8 as they're the only ones
# with persistence data. Plus for some reason
# the persistence replicates are numbered from 8
persist.dat = my.dat %>% filter(
      Substrate != 'Cott01' & 
      Experiment %in% c(7,8) &
      Replicate > 7 &
      ObservationType != 'Ndata') %>%
      select(c("Substrate", "ObservationType", "Count", "Mass", "PersistenceTime", "Experiment", "Replicate", "Note"))

# fix column types
persist.dat$Count = as.numeric(persist.dat$Count)
persist.dat$PersistenceTime = as.numeric(persist.dat$PersistenceTime)
persist.dat$Experiment = as.character(persist.dat$Experiment)
persist.dat$Replicate = as.character(persist.dat$Replicate)

# summarise the count data
summ.dat = summarySE(persist.dat, measurevar = "Count", groupvars = c("Substrate", "PersistenceTime", "Experiment", "Note"))
knitr::kable(summ.dat, caption = "Summary of EM's Persistence Data")

```

However, there are nowhere near enough replicates to do anything with the data. Only have 2 reps for 0, 30, 60 and 1440 minutes. All the rest are single-point. #sadface

The difference between experiments '7' and '8' are the camera settings 'C1' and 'C2' which, if they weren't, too different could combine to get more replicates. However, the 'C2' counts are consistently higher than 'C1' suggesting that they're not the same and cannot be considered replicates.

```{r emplot}

summ.dat = persist.dat %>% 
  group_by(Substrate, PersistenceTime, Note) %>%
  summarise(N = n(), Mean = mean(Count), Max = max(Count), Min = min(Count))

ggplot(summ.dat, aes(x=PersistenceTime, y = Mean, colour = Note)) +
  geom_point() +
  geom_linerange(aes(ymin = Min, ymax = Max)) +
  xlim(c(-10, 1500)) +
  labs(title = "Nylon Persistence",
       subtitle = "By different camera settings",
       x = "Persistence Time (min)",
       y = "Particle Count",
       caption = "Error bars: min-max") +
  mytheme
  
```
Yup. Data are all over the place. Can't use this in the paper.
