---
title: "A Closer Look at Particle Counts"
date: "30/06/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, tidy = TRUE)

library(ggplot2)
library(dplyr)
```

## Background

Expt D in the Transfer studies shows high variability and much higher transfer ratio, $T_r$, values than the others, especially with wool. On closer inspection it appears that it (as well as expt B to some extent) shows quite high values for Photo 1, $P_1$, - median ~100) - when the A,C & E are much near to zero.

With that information can simulate or reproduce the behaviour on simulated data. As all the data are raw counts of independent events it is reasonable to assume a Poisson distribution. Poisson distributions are dependent on two parameters $k$, the number of occurrences, and $\mathbf{\lambda}$, the mean (which is also equal to the variance). 

The hypothesis being are relatively high $P_1$ counts causative of high Transfer Ratios independently of other photo counts?

So we can model the counts for photos $P_1$, $P_2$ and $P_3$ assuming a Poisson distribution and taking equivalent means from the Transfer experiment. Photos $P_4$ and $P_5$ are a function of these starting counts based on the amount of transfer. A 35% transfer of particles is consistent with the data. A random normal sampling around this mean with a standard deviation of 4 was used.

$n = 20$ for the number of times each transfer 'experiment' was simulated.

Loss and 'declumping' of particles was not simulated.

## Simulation


```{r func}
# function to produce the starting data
makeData <- function(p1 = 45, p2 = 6, p3 = 120, n = 20, transfer.pc = 35) {
  # create donor background data
  p1.dist = rpois(n, p1)
  
  # low background on receiver
  p2.dist = rpois(n, p2)
  # add some particles
  p3.dist = rpois(n, p3)
  
  # assume 30-40 percent transfer
  trans = round(rnorm(n, transfer.pc, sd = 4))/100
  
  # create modelled data for particulate transfer
  df = data.frame(p1.dist, p2.dist, p3.dist, p4= p1.dist + round(p3.dist*trans), p5 = p1.dist + round(p3.dist*(1-trans)))
  
  return(df)
}
```

With the function for creating the distributions defined a range of counts was parameterised:

  * $P_1$ = 5, 30, 55, 80, 105
  * $P_2$ = 5, 30, 55, 80, 105
  * $P_3$ = 100, 150, 200, 250, 300


```{r data}

df.dat = data.frame()
for(p1 in seq(5,105,25)) {
  for(p2 in seq(5,105,25)) {
    for(p3 in seq(100,300,50)) {
      df = makeData(p1, p2, p3)
      if (nrow(df.dat) > 0) {
        tmp = data.frame(p1.m = as.character(p1), p2.m = as.character(p2), p3.m = as.character(p3), df)
        df.dat = rbind(df.dat, tmp)
      } else {
        df.dat = data.frame(p1.m = as.character(p1), p2.m = as.character(p2), p3.m = as.character(p3), df)
      }
    }
  }
}
```

The Transfer Ratio was calculated as previous described:

$$
  t_r = (P_5 - P_2) / (P_3 - P_1)
$$


```{r tr}

# calc transfer ratio
df.dat$tr = (df.dat$p5 - df.dat$p2.dist)/(df.dat$p3.dist - df.dat$p1.dist)

```

**Note**: If $P_1$ = $P_3$ the denominator is $0$ which results in a non-finite $T_r$.

## Results

If we fix $P_2 = 5$ and simulate $P_1$ and $P_3$ we can visualise the effect on $T_r$ as a function of both counts.

```{r p1plt, fig.cap='Simulation of Transfer Ratio as a Function of P1 & P3 Counts'}
df.dat %>% filter(p2.m == '5') %>%
ggplot(aes(x = factor(p1.m, levels = c(seq(5,105,25))), y = tr)) + 
  labs(title = "Simulation of Transfer Ratio as a Function of P1 & P3 Counts",
       subtitle = 'n = 20, Mean P2 counts = 5, % particles transferred = 35% (+/- 4 s.d.)',
       x = 'Mean P1 Counts',
       y = 'Transfer Ratio',
       caption = 'Y-axes are not the same') +
  geom_boxplot() +
  facet_wrap(p3.m ~ ., scales = 'free_y')

ggsave('p1_count_simulation.png')

```

Figure 1 shows that for $P_3 > 150$ there is a predictable, linear relationship between $P_1$ and $P_3$ counts. However, with $P_3 \leq 150$  and mean $P_1$ increasing the $T_r$ relationship gets very unstable - high positive and negative values. This suggests that it's not just high $P_1$ counts that are the problem, but it's when $P_1 \approx P_3$ then $T_r$ is extreme.

As a comparison if we look at the $P_2$, the receiver background, relationship with $T_r$ there is no issue either at high counts or when $P_2 \approx P_3$.

```{r p2plt, fig.cap = 'Simulation of Transfer Ratio as a Function of P2 Counts'}
df.dat %>% filter(p1.m == 5) %>%
ggplot( aes(x = factor(p2.m, levels = c(seq(5,105,25))), y = tr)) + 
  labs(title = 'Simulation of Transfer Ratio as a Function of P2 Counts',
       subtitle = 'n = 20, Mean P1 counts = 5, % particles transferred = 35% (+/- 4 s.d.)',
       x = 'Mean P2 Counts',
       y = 'Transfer Ratio',
       caption = 'Y-axes are the same') +
  geom_boxplot() +
  facet_wrap(p3.m ~ .)

```

## Conclusion

High $P_1$ - background receiver counts - are not the cause of the observed
issues with $T_r$ in expt D, but it appears to be when $P_1 \approx P_3$ that
results in very high or negative $T_r$ values. This is not the case for expt D
which suggests that the hypothesis that high $P_1$ counts are causative of high
Transfer Ratios is incorrect. 




