---
title: "Transfer Data Analysis"
author: "Christian Cole (original author), Hervé Ménard (code editing), Niamh Nic Daéid (code editing)"
output:  pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
# libraries
library(readxl)
library(plyr)
library(ggplot2)
library(ggpubr)
library(forcats)
library(tibble)
library(tidyr)
source('lib.R')
library(dplyr)
```

```{r themes}

mytheme = theme_minimal() +
  theme(panel.grid.minor = element_blank(),
        panel.grid.major.x = element_blank(),
        axis.title = element_text(size = 13, face = "bold"),
        axis.text = element_text(size = 11),
        legend.title = element_blank())


```

# Dataset A


```{r readA}
# Read in A
my.dat = read_xlsx("dat/Counts_data.xlsx", sheet = "A")

# only use the transfer experiments 1 & 2. Experiments 3-6 are not 
# clear in what they are and don't fit with the rest. Mostly Persistence?
transfer.dat = my.dat %>% filter(`PersistenceTime (min)` == 0 & Experiment %in% c(1,2)) %>%
  select(c("Substrate", "ObservationType", "Count", "Mass (g)", "TransferTime (s)", "Experiment", "Replicate")) %>%
  dplyr::rename(Mass = `Mass (g)`, 
         TransferTime = `TransferTime (s)`)

knitr::kable(head(transfer.dat, 20), caption = "'Head' (i.e. first 20 rows) of A Transfer Data")

```


```{r mungeA}

# fix column types
transfer.dat$Count = as.numeric(transfer.dat$Count)
transfer.dat$Experiment = as.character(transfer.dat$Experiment)
transfer.dat$Replicate = as.character(transfer.dat$Replicate)

# add a photoID column to keep track of which datapoint is which
photo.id = rep(paste0("P", seq(1:5)), as.integer(nrow(transfer.dat)/5+1))
transfer.dat = data.frame(PhotoID = photo.id[1:nrow(transfer.dat)], transfer.dat)

# split out each set of photos...
p1 = transfer.dat %>% filter(PhotoID == "P1") %>% select(Mass,TransferTime,Substrate,Experiment,Replicate,Count)
p2 = transfer.dat %>% filter(PhotoID == "P2") %>% select(Mass,TransferTime,Substrate,Experiment,Replicate,Count)
p3 = transfer.dat %>% filter(PhotoID == "P3") %>% select(Mass,TransferTime,Substrate,Experiment,Replicate,Count)
p4 = transfer.dat %>% filter(PhotoID == "P4") %>% select(Mass,TransferTime,Substrate,Experiment,Replicate,Count)
p5 = transfer.dat %>% filter(PhotoID == "P5") %>% select(Mass,TransferTime,Substrate,Experiment,Replicate,Count)

# ...and merge
merge.dat = data.frame(p5[,c("Mass","TransferTime","Substrate","Experiment","Replicate")], p1$Count, p2$Count, p3$Count, p4$Count, p5$Count)
names(merge.dat) = c("Mass","TransferTime","Substrate","Experiment","Replicate","P1","P2","P3","P4","P5")

all.dat = cbind(Set = rep("A",nrow(merge.dat)), merge.dat)


# calculate the Transfer Ratios and Efficiencies
# NB: the summarise() function can be problematic here, if plyr is loaded
#     after dplyr. Check that first if this stops working.
final.dat = merge.dat %>% group_by(Mass, TransferTime, Experiment, Replicate) %>% mutate(Donor = P3 - P1, Receiver = P5 - P2, Donor_post = P3 - P4) %>% mutate(Ratio = Receiver/Donor, Efficiency = Receiver/Donor_post)

summ.dat = summarySE(final.dat, measurevar = "Ratio", groupvars = c("Experiment"))
ratio.dat = summ.dat[, -1]
ratio.dat = cbind(Mass = rep(1000,2), TransferTime = rep(60,2), Substrate = c("Wool01", "Nylo01"), ratio.dat)


knitr::kable(ratio.dat, caption = "Summary Data for A Transfer Ratios")
```

```{r plotA, fig.height= 4}

ggplot(summ.dat, aes(x = Experiment, y = Ratio)) + 
  geom_point(size=2.2) +
  geom_errorbar(aes(ymin = Ratio-se, ymax = Ratio+se), width = 0.08) +
  geom_point(colour = "white") +
  scale_x_discrete(labels = c("Wool", "Nylon")) +
  labs(title = "Analysis of Transfer Ratios",
       x = "Receiver Material",
       y = "Transfer Ratio")
```

Figure S1: Transfer Ratios Dataset A

```{r ploteA, fig.height= 4}

eff.dat = summarySE(final.dat, measurevar = "Efficiency", groupvars = c("Experiment"))

#eff.dat
ggplot(eff.dat, aes(x = Experiment, y = Efficiency)) + 
  geom_point(size=1) +
  geom_errorbar(aes(ymin = Efficiency-se, ymax = Efficiency+se), width = 0.08) +
  geom_point(colour = "white", size = 0.7) +
  scale_x_discrete(labels = c("Wool", "Nylon")) +
  labs(title = "Analysis of Transfer Efficiency",
       x = "",
       y = "Efficiency") 

```

Figure S2: Transfer Efficiency dataset A.

```{r alldat, eval=FALSE}

# start df of all ratio data for comparison
all.ratio = summ.dat
all.ratio = all.ratio[, -1]
all.ratio = cbind(Set = rep("A", 2), Mass = rep(1000,2), TransferTime = rep(60,2), Substrate = c("Wool01", "Nylo01"), all.ratio)

# do same for efficiency data
all.eff = eff.dat
all.eff = all.eff[, -1]
all.eff = cbind(Set = rep("A", 2), Mass = rep(1000,2), TransferTime = rep(60,2), Substrate = c("Wool01", "Nylo01"), all.eff)

```

# Dataset B


```{r readB}
# Read in B
my.dat = read_xlsx("dat/Counts_data.xlsx", sheet = "B", trim_ws = TRUE)

# only use the transfer experiments for Groups 8-22. Groups 1-7 have no 
# count data and group 23 is incomplete, only one replicate
transfer.dat = my.dat %>% filter(`PersistenceTime (min)` == 0 & Experiment > 8 & Experiment < 23) %>%
  select(c("Substrate", "ObservationType", "Count", "Mass (g)", "TransferTime (s)", "Experiment", "Replicate")) %>%
  dplyr::rename(Mass = `Mass (g)`, 
         TransferTime = `TransferTime (s)`)

knitr::kable(head(transfer.dat, 20), caption = "'Head' of B Transfer Data")

```
```{r mungeB}

# fix column types
transfer.dat$Count = as.numeric(transfer.dat$Count)
transfer.dat$TransferTime = as.numeric(transfer.dat$TransferTime)
transfer.dat$Mass = as.numeric(transfer.dat$Mass)
transfer.dat$Experiment = as.character(transfer.dat$Experiment)
transfer.dat$Replicate = as.character(transfer.dat$Replicate)

# rename Denm01 to Deni01 to match others
transfer.dat[transfer.dat$Substrate == "Denm01", 1] <- "Deni01"

# add a photoID column to keep track of which datapoint is which
photo.id = rep(paste0("P", seq(1:5)), as.integer(nrow(transfer.dat)/5+1))
transfer.dat = data.frame(PhotoID = photo.id[1:nrow(transfer.dat)], transfer.dat)

# split out each set of photos...
p1 = transfer.dat %>% filter(PhotoID == "P1") %>% select(Mass,TransferTime,Substrate,Experiment,Replicate,Count)
p2 = transfer.dat %>% filter(PhotoID == "P2") %>% select(Mass,TransferTime,Substrate,Experiment,Replicate,Count)
p3 = transfer.dat %>% filter(PhotoID == "P3") %>% select(Mass,TransferTime,Substrate,Experiment,Replicate,Count)
p4 = transfer.dat %>% filter(PhotoID == "P4") %>% select(Mass,TransferTime,Substrate,Experiment,Replicate,Count)
p5 = transfer.dat %>% filter(PhotoID == "P5") %>% select(Mass,TransferTime,Substrate,Experiment,Replicate,Count)

# ...and merge
merge.dat = data.frame(p5[,c("Mass","TransferTime","Substrate","Experiment","Replicate")], p1$Count, p2$Count, p3$Count, p4$Count, p5$Count)
names(merge.dat) = c("Mass","TransferTime","Substrate","Experiment","Replicate","P1","P2","P3","P4","P5")

all.dat = rbind(all.dat, data.frame(Set = rep("B",nrow(merge.dat)), merge.dat))

# calculate the Transfer Ratios and Efficiencies
# NB: the summarise() function can be problematic here, if plyr is loaded
#     after dplyr. Check that first if this stops working.
final.dat = merge.dat %>% group_by(Mass, TransferTime, Substrate, Replicate) %>% mutate(Donor = P3 - P1, Receiver = P5 - P2, Donor_post = P3 - P4) %>% mutate(Ratio = Receiver/Donor, Efficiency = Receiver/Donor_post)

ratio.dat = summarySE(final.dat, measurevar = "Ratio", groupvars = c("Mass", "TransferTime", "Substrate"))

knitr::kable(ratio.dat, caption = "Summary Data for B Transfer Ratios")
```

```{r plottB}

ggplot(ratio.dat, aes(x = numericFactorOrder(TransferTime), y = Ratio)) + 
  geom_point(size=2.2) +
  geom_errorbar(aes(ymin = Ratio-se, ymax = Ratio+se), width = 0.08) +
  geom_point(colour = "white") +
  labs(title = "Analysis of Transfer Ratios",
       x = "Receiver Material",
       y = "Transfer Ratio") +
  facet_grid(numericFactorOrder(Mass) ~ Substrate)


```

Figure S3: Transfer Ratios dataset B.

```{r ploteB}

eff.dat = summarySE(final.dat, measurevar = "Efficiency", groupvars = c("Mass","TransferTime","Substrate"))
knitr::kable(eff.dat, caption = "Efficiency data for B")

ggplot(eff.dat, aes(x = numericFactorOrder(TransferTime), y = Efficiency)) + 
  geom_point(size=1) +
  geom_errorbar(aes(ymin = Efficiency-se, ymax = Efficiency+se), width = 0.08) +
  geom_point(colour = "white", size = 0.7) +
#  scale_x_discrete(labels = c("Wool", "Nylon")) +
  labs(title = "Analysis of Transfer Efficiency",
       x = "Transfer Time (s)",
       y = "Efficiency") +
  facet_grid(numericFactorOrder(Mass) ~ Substrate)


```

Figure S4: Transfer Efficiency dataset B.

\newpage
# Dataset C


```{r readC}
# Read in C data
my.dat = read_xlsx("dat/Counts_data.xlsx", sheet = "C", trim_ws = TRUE, col_types = "text")

# only use the Transfer data (PersistenceTime == "0") 
transfer.dat = my.dat %>% filter(`PersistenceTime (min)` == 0) %>%
  select(c("Substrate", "ObservationType", "Count", "Mass (g)", "TransferTime (s)", "Experiment", "Replicate")) %>%
  dplyr::rename(Mass = `Mass (g)`, 
         TransferTime = `TransferTime (s)`)

knitr::kable(head(transfer.dat,20), caption = "'Head' of C Transfer Data")

```
```{r mungeC}
# fix column types
transfer.dat$Count = as.numeric(transfer.dat$Count)
transfer.dat$Experiment = as.character(transfer.dat$Experiment)
transfer.dat$Replicate = as.character(transfer.dat$Replicate)

# add a photoID column to keep track of which datapoint is which
photo.id = rep(paste0("P", seq(1:5)), as.integer(nrow(transfer.dat)/5+1))
transfer.dat = data.frame(PhotoID = photo.id[1:nrow(transfer.dat)], transfer.dat)

# split out each set of photos...
p1 = transfer.dat %>% filter(PhotoID == "P1") %>% select(Mass,TransferTime,Substrate,Experiment,Replicate,Count)
p2 = transfer.dat %>% filter(PhotoID == "P2") %>% select(Mass,TransferTime,Substrate,Experiment,Replicate,Count)
p3 = transfer.dat %>% filter(PhotoID == "P3") %>% select(Mass,TransferTime,Substrate,Experiment,Replicate,Count)
p4 = transfer.dat %>% filter(PhotoID == "P4") %>% select(Mass,TransferTime,Substrate,Experiment,Replicate,Count)
p5 = transfer.dat %>% filter(PhotoID == "P5") %>% select(Mass,TransferTime,Substrate,Experiment,Replicate,Count)

# ...and merge
merge.dat = data.frame(p5[,c("Mass","TransferTime","Substrate","Experiment","Replicate")], p1$Count, p2$Count, p3$Count, p4$Count, p5$Count)
names(merge.dat) = c("Mass","TransferTime","Substrate","Experiment","Replicate","P1","P2","P3","P4","P5")

all.dat = rbind(all.dat, data.frame(Set = rep("C",nrow(merge.dat)), merge.dat))

# calculate the Transfer Ratios and Efficiencies
# NB: the summarise() function can be problematic here, if plyr is loaded
#     after dplyr. Check that first if this stops working.
final.dat = merge.dat %>% group_by(Mass, TransferTime, Substrate, Replicate) %>% mutate(Donor = P3 - P1, Receiver = P5 - P2, Donor_post = P3 - P4) %>% mutate(Ratio = Receiver/Donor, Efficiency = Receiver/Donor_post)

ratio.dat = summarySE(final.dat, measurevar = "Ratio", groupvars = c("Mass","TransferTime","Substrate"))

knitr::kable(ratio.dat, caption = "Summary Data for C Transfer Ratios")
```

```{r plottC}

ggplot(ratio.dat, aes(x = numericFactorOrder(TransferTime), y = Ratio)) + 
  geom_point(size=1) +
  geom_errorbar(aes(ymin = Ratio-se, ymax = Ratio+se), width = 0.08) +
  geom_point(colour = "white", size = 0.7) +
#  scale_x_discrete(labels = c("Wool", "Nylon")) +
  labs(title = "Analysis of Transfer Ratio",
       x = "Transfer Time (s)",
       y = "Transfer Ratio") +
  facet_grid(numericFactorOrder(Mass) ~ Substrate)


```

Figure S5: Transfer Ratio dataset C.

```{r ploteC}

eff.dat = summarySE(final.dat, measurevar = "Efficiency", groupvars = c("Mass","TransferTime","Substrate"))

knitr::kable(eff.dat, caption = "Efficiency data for C")

ggplot(eff.dat, aes(x = numericFactorOrder(TransferTime), y = Efficiency)) + 
  geom_point(size=1) +
  geom_errorbar(aes(ymin = Efficiency-se, ymax = Efficiency+se), width = 0.08) +
  geom_point(colour = "white", size = 0.7) +
#  scale_x_discrete(labels = c("Wool", "Nylon")) +
  labs(title = "Analysis of Transfer Efficiency",
       x = "Transfer Time (s)",
       y = "Efficiency") +
  facet_grid(numericFactorOrder(Mass) ~ Substrate)


```

Figure S6: Transfer Efficiency dataset C.

\newpage
## Dataset D


```{r readD}
# Read in D data
my.dat = read_xlsx("dat/Counts_data.xlsx", sheet = "D", trim_ws = TRUE, col_types = "text")

# only use the Transfer data (PersistenceTime == "0") 
transfer.dat = my.dat %>% filter(`PersistenceTime (min)` == 0) %>%
  select(c("Substrate", "ObservationType", "Count", "Mass (g)", "TransferTime (s)", "Experiment", "Replicate")) %>%
  dplyr::rename(Mass = `Mass (g)`, 
         TransferTime = `TransferTime (s)`)

# also remove expt 16 as it is incomplete
transfer.dat = transfer.dat[as.numeric(transfer.dat$Experiment) < 16,]

knitr::kable(head(transfer.dat,20), caption = "'Head' of D Transfer Data")

```


```{r mungeD}
# fix column types
transfer.dat$Count = as.numeric(transfer.dat$Count)
transfer.dat$Experiment = as.character(transfer.dat$Experiment)
transfer.dat$Replicate = as.character(transfer.dat$Replicate)

# add a photoID column to keep track of which datapoint is which
photo.id = rep(paste0("P", seq(1:5)), as.integer(nrow(transfer.dat)/5+1))
transfer.dat = data.frame(PhotoID = photo.id[1:nrow(transfer.dat)], transfer.dat)

# split out each set of photos...
p1 = transfer.dat %>% filter(PhotoID == "P1") %>% select(Mass,TransferTime,Substrate,Experiment,Replicate,Count)
p2 = transfer.dat %>% filter(PhotoID == "P2") %>% select(Mass,TransferTime,Substrate,Experiment,Replicate,Count)
p3 = transfer.dat %>% filter(PhotoID == "P3") %>% select(Mass,TransferTime,Substrate,Experiment,Replicate,Count)
p4 = transfer.dat %>% filter(PhotoID == "P4") %>% select(Mass,TransferTime,Substrate,Experiment,Replicate,Count)
p5 = transfer.dat %>% filter(PhotoID == "P5") %>% select(Mass,TransferTime,Substrate,Experiment,Replicate,Count)

# ...and merge
merge.dat = data.frame(p5[,c("Mass","TransferTime","Substrate","Experiment","Replicate")], p1$Count, p2$Count, p3$Count, p4$Count, p5$Count)
names(merge.dat) = c("Mass","TransferTime","Substrate","Experiment","Replicate","P1","P2","P3","P4","P5")

all.dat = rbind(all.dat, data.frame(Set = rep("D",nrow(merge.dat)), merge.dat))

# calculate the Transfer Ratios and Efficiencies
# NB: the summarise() function can be problematic here, if plyr is loaded
#     after dplyr. Check that first if this stops working.
final.dat = merge.dat %>% group_by(Mass, TransferTime, Substrate, Replicate) %>% mutate(Donor = P3 - P1, Receiver = P5 - P2, Donor_post = P3 - P4) %>% mutate(Ratio = Receiver/Donor, Efficiency = Receiver/Donor_post)

ratio.dat = summarySE(final.dat, measurevar = "Ratio", groupvars = c("Mass","TransferTime","Substrate"))

knitr::kable(ratio.dat, caption = "Summary Data for D Transfer Ratios")
```

```{r plottD}

ggplot(ratio.dat, aes(x = numericFactorOrder(TransferTime), y = Ratio)) + 
  geom_point(size=1) +
  geom_errorbar(aes(ymin = Ratio-se, ymax = Ratio+se), width = 0.08) +
  geom_point(colour = "white", size = 0.7) +
#  scale_x_discrete(labels = c("Wool", "Nylon")) +
  labs(title = "Analysis of Transfer Ratio",
       x = "Transfer Time (s)",
       y = "Transfer Ratio") +
  facet_grid(numericFactorOrder(Mass) ~ Substrate)


```

Figure S7: Transfer Ratio dataset D.

```{r ploteD}

eff.dat = summarySE(final.dat, measurevar = "Efficiency", groupvars = c("Mass","TransferTime","Substrate"))
eff.dat = cbind(Set = rep("D", nrow(eff.dat)), eff.dat)

knitr::kable(eff.dat, caption = "Efficiency data for D")

ggplot(eff.dat, aes(x = numericFactorOrder(TransferTime), y = Efficiency)) + 
  geom_point(size=1) +
  geom_errorbar(aes(ymin = Efficiency-se, ymax = Efficiency+se), width = 0.08) +
  geom_point(colour = "white", size = 0.7) +
#  scale_x_discrete(labels = c("Wool", "Nylon")) +
  labs(title = "Analysis of Transfer Efficiency",
       x = "Transfer Time (s)",
       y = "Efficiency") +
  facet_grid(numericFactorOrder(Mass) ~ Substrate)


```

Figure S8: Transfer Efficiency dataset D.

\newpage
## Dataset E


```{r readE}
# Read in E data
my.dat = read_xlsx("dat/Counts_data.xlsx", sheet = "E", trim_ws = TRUE, col_types = "text")

# only use the Transfer data (PersistenceTime == "0") 
transfer.dat = my.dat %>% filter(`PersistenceTime (min)` == 0) %>%
  select(c("Substrate", "ObservationType", "Count", "Mass (g)", "TransferTime (s)", "Experiment", "Replicate", "Note")) %>%
  dplyr::rename(Mass = `Mass (g)`, 
         TransferTime = `TransferTime (s)`)

# also remove expt 16 as it is incomplete
#transfer.dat = transfer.dat[as.numeric(transfer.dat$Experiment) < 16,]

knitr::kable(head(transfer.dat, 20), caption = "'Head' of E Transfer Data")

```

\newpage
```{r mungeE}
# fix column types
transfer.dat$Count = as.numeric(transfer.dat$Count)
transfer.dat$Experiment = as.character(transfer.dat$Experiment)
transfer.dat$Replicate = as.character(transfer.dat$Replicate)

# add a photoID column to keep track of which datapoint is which
photo.id = rep(paste0("P", seq(1:5)), as.integer(nrow(transfer.dat)/5+1))
transfer.dat = data.frame(PhotoID = photo.id[1:nrow(transfer.dat)], transfer.dat)

# split out each set of photos...
p1 = transfer.dat %>% filter(PhotoID == "P1") %>% select(Mass,TransferTime,Substrate,Experiment,Replicate,Note,Count)
p2 = transfer.dat %>% filter(PhotoID == "P2") %>% select(Mass,TransferTime,Substrate,Experiment,Replicate,Note,Count)
p3 = transfer.dat %>% filter(PhotoID == "P3") %>% select(Mass,TransferTime,Substrate,Experiment,Replicate,Note,Count)
p4 = transfer.dat %>% filter(PhotoID == "P4") %>% select(Mass,TransferTime,Substrate,Experiment,Replicate,Note,Count)
p5 = transfer.dat %>% filter(PhotoID == "P5") %>% select(Mass,TransferTime,Substrate,Experiment,Replicate,Note,Count)

# ...and merge
merge.dat = data.frame(p5[,c("Mass","TransferTime","Substrate","Experiment","Replicate", "Note")], p1$Count, p2$Count, p3$Count, p4$Count, p5$Count)
names(merge.dat) = c("Mass","TransferTime","Substrate","Experiment","Replicate", "Note","P1","P2","P3","P4","P5")

# this is a little more complicated as need select only C1
# data and then remove the note column before combining
# with the other datasets
e.dat = merge.dat[merge.dat$Note == 'C1',]
e.dat = e.dat[, -6]
all.dat = rbind(all.dat, data.frame(Set = rep("E",nrow(e.dat)), e.dat))

# calculate the Transfer Ratios and Efficiencies
# NB: the summarise() function can be problematic here, if plyr is loaded
#     after dplyr. Check that first if this stops working.
final.dat = merge.dat %>% group_by(Mass, TransferTime, Experiment, Substrate, Replicate, Note) %>% mutate(Donor = P3 - P1, Receiver = P5 - P2, Donor_post = P3 - P4) %>% mutate(Ratio = Receiver/Donor, Efficiency = Receiver/Donor_post)

ratio.dat = summarySE(final.dat, measurevar = "Ratio", groupvars = c("Mass","TransferTime","Substrate", "Experiment", "Note"))
# remove experimental conditions with < 2 replicates.
ratio.dat = ratio.dat[ratio.dat$N > 1,]

knitr::kable(ratio.dat, caption = "Summary Data for E Transfer Ratios")
```

```{r plottE, fig.height= 8}

tc1 = ggplot(ratio.dat[ratio.dat$Note == 'C1',], aes(x = numericFactorOrder(TransferTime), y = Ratio)) + 
  geom_point(size=1) +
  geom_errorbar(aes(ymin = Ratio-se, ymax = Ratio+se), width = 0.08) +
  geom_point(colour = "white", size = 0.7) +
#  scale_x_discrete(labels = c("Wool", "Nylon")) +
  labs(title = "Analysis of Transfer Ratio - C1",
       x = "Transfer Time (s)",
       y = "Transfer Ratio") +
  facet_grid(numericFactorOrder(Mass) ~ Substrate)

tc2 = ggplot(ratio.dat[ratio.dat$Note == 'C2',], aes(x = numericFactorOrder(TransferTime), y = Ratio)) + 
  geom_point(size=1) +
  geom_errorbar(aes(ymin = Ratio-se, ymax = Ratio+se), width = 0.08) +
  geom_point(colour = "white", size = 0.7) +
#  scale_x_discrete(labels = c("Wool", "Nylon")) +
  labs(title = "Analysis of Transfer Ratio - C2",
       x = "Transfer Time (s)",
       y = "Transfer Ratio") +
  facet_grid(numericFactorOrder(Mass) ~ Substrate)

tc12 = ggarrange(tc1, tc2, labels = c('A', 'B'), nrow = 2,
                  common.legend = TRUE)

tc12
```
Figure S9: Transfer Ratio dataset E. (A) Camera setting C1, and (B) Camera setting C2.
\newpage
```{r ploteF}

eff.dat = summarySE(final.dat, measurevar = "Efficiency", groupvars = c("Mass","TransferTime","Substrate"))

knitr::kable(eff.dat, caption = "Efficiency data for E")
```


```{r ploteE}
ggplot(eff.dat, aes(x = numericFactorOrder(TransferTime), y = Efficiency)) + 
  geom_point(size=1) +
  geom_errorbar(aes(ymin = Efficiency-se, ymax = Efficiency+se), width = 0.08) +
  geom_point(colour = "white", size = 0.7) +
#  scale_x_discrete(labels = c("Wool", "Nylon")) +
  labs(title = "Analysis of Transfer Efficiency",
       x = "Transfer Time (s)",
       y = "Efficiency") +
  facet_grid(numericFactorOrder(Mass) ~ Substrate)


```

Figure S10: Transfer Efficiency dataset E.

\newpage
# Combined Analysis - counts


```{r countcomp1}

# 
all_counts.dat = all.dat %>% 
  gather(photo, count, P1:P5) 

# remove '0' padding in Transfer time
all_counts.dat$TransferTime = as.numeric(all_counts.dat$TransferTime)

# %>% 
#   group_by(Set, photo) %>% 
#   summarise(N = n(), mean = mean(count), SD = sd(count), SE = sd(count)/sqrt(n())) 
knitr::kable(head(all_counts.dat, 10), caption = "Summary statistics of particle counts for each photo type by experiment")

```

```{r countcomp2,fig.height=3.9}
ggplot(all_counts.dat, aes(x = Set, y = count)) + 
  labs(title = 'Particle Counts Per Photo',
       subtitle = 'All materials, times and masses',
       x = 'Transfer Experiment',
       y = 'Particle Count') +
  geom_boxplot(width = 0.5)+
  facet_wrap(photo ~ .)

#ggsave('cnt1.png')
```

Figure S11: Particle Counts Per Photo, all datasets, materials, masses and times.

```{r countcomp3,fig.height=3.94}
all_counts.dat %>% filter(Substrate == 'Wool01') %>%
ggplot(aes(x = Set, y = count)) + 
  labs(title = 'Particle Counts Per Photo',
       subtitle = 'Wool and all times and masses',
       x = 'Transfer Experiment',
       y = 'Particle Count') +
  geom_boxplot(width = 0.5) +
  facet_wrap(photo ~ .)

#ggsave('cnt2.png')

```

Figure S12: Particle counts per photo, Wool, all datasets, masses and times.

```{r countcomp4,fig.height= 3.9}
all_counts.dat %>% filter(Substrate == 'Nylo01') %>%
ggplot(aes(x = Set, y = count)) + 
  labs(title = 'Particle Counts Per Photo',
       subtitle = 'Nylon and all times and masses',
       x = 'Transfer Experiment',
       y = 'Particle Count') +
  geom_boxplot(width = 0.5) +
  facet_wrap(photo ~ .)

#ggsave('cnt3.png')

```

Figure S13: Particle counts per photo, Nylon, all datasets, masses and times.

```{r countcomp5,fig.height= 3.9}
all_counts.dat %>% filter(Substrate == 'Wool01', Mass == '1000', TransferTime == '60') %>%
ggplot(aes(x = Set, y = count)) + 
  labs(title = 'Particle Counts Per Photo',
       subtitle = 'Wool, 60 seconds, 1000 grams',
       x = 'Transfer Experiment',
       y = 'Particle Count') +
  geom_boxplot(width = 0.5) +
  facet_wrap(photo ~ .)

#ggsave('cnt4.png')

```

Figure S14: Particle counts per photo, Wool, all datasets, 60 s, 1000 g.

```{r countcomp6,fig.height= 3.9}
all_counts.dat %>% filter(Substrate == 'Nylo01', Mass == '1000', TransferTime == '60') %>%
ggplot(aes(x = Set, y = count)) + 
  labs(title = 'Particle Counts Per Photo',
       subtitle = 'Nylon, 60 seconds, 1000 grams',
       x = 'Transfer Experiment',
       y = 'Particle Count') +
  geom_boxplot(width = 0.5) +
  facet_wrap(photo ~ .)

#ggsave('cnt5.png')

```

Figure S15: Particle counts per photo, Nylon, all datasets, 60 s, 1000 g.

```{r p3minusp1,fig.height= 3.9}

# redo long format now including the P3-P1 difference 
all_counts.dat = all.dat %>% 
  mutate(diffp3p1 = P3 - P1) %>%
  gather(photo, count, P1:diffp3p1) 

# remove '0' padding in Transfer time
all_counts.dat$TransferTime = as.numeric(all_counts.dat$TransferTime)


labels = list(
  'diffp3p1' = 'P3-P1',
  'P1' = 'Photo 1',
  'P2' = 'Photo 2',
  'P3' = 'Photo 3',
  'P4' = 'Photo 4',
  'P5' = 'Photo 5'
)
photo_labeller <- function(variable,value){
  return(labels[value])
}

all_counts.dat %>% 
#  filter(Substrate == 'Wool01',Mass == '1000', TransferTime == '60') %>%
  ggplot(aes(x = Set, y = count)) +
  labs(title = 'Particle Counts Per Photo',
       subtitle = 'All materials, times and masses',
       x = 'Transfer Experiment',
       y = 'Particle Count') +
  geom_boxplot(width = 0.5)+
  geom_boxplot() +
  facet_wrap(photo ~ ., labeller = photo_labeller)

```

Figure S16: Particle counts per photo, all datasets, materials, masses and times. Included top left: Donor after UV deposition (Photo 3) - Donor Background (Photo 1).

```{r,fig.height= 3.9}

all_counts.dat %>% 
  filter(Substrate == 'Wool01') %>%
  ggplot(aes(x = Set, y = count)) +
  labs(title = 'Particle Counts Per Photo',
       subtitle = 'Wool, all times and masses',
       x = 'Transfer Experiment',
       y = 'Particle Count') +
  geom_boxplot(width = 0.5)+
  geom_boxplot() +
  facet_wrap(photo ~ ., labeller = photo_labeller)


```

Figure S17: Particle counts per photo, Wool, all datasets, masses and time. Included top left: Donor after UV deposition (Photo 3) - Donor Background (Photo 1).

\newpage
# Combined Analysis - ratios


```{r plottr1,fig.height= 3.9}

all.dat$Mass = as.numeric(all.dat$Mass)
all.dat$TransferTime = as.numeric(all.dat$TransferTime)
all.dat[all.dat$Substrate == 'Wool01', 'Substrate'] <- 'Wool'
all.dat[all.dat$Substrate == 'Nylo01', 'Substrate'] <- 'Nylon'
all.dat[all.dat$Substrate == 'Elas01', 'Substrate'] <- 'Elastane'
all.dat[all.dat$Substrate == 'Deni01', 'Substrate'] <- 'Denim'

# calc transfer ratios and efficiencies
all.final = all.dat %>% group_by(Set, Mass, TransferTime, Experiment, Substrate, Replicate) %>%
  summarise(Donor = P3 - P1, Receiver = P5 - P2, Donor_post = P3 - P4) %>% mutate(Ratio = Receiver/Donor, Efficiency = Receiver/Donor_post)

# aggregate over replicates
all.ratio = all.final %>% 
  group_by(Set, Mass, TransferTime, Substrate) %>%
  summarise(N = n(), ratio = mean(Ratio), se = stdErr(Ratio))

m1000.dat = all.ratio %>% filter(Mass == 1000 & Substrate %in% c('Wool', 'Nylon'))
#wool.dat

trp1 = ggplot(m1000.dat[m1000.dat$TransferTime != '10',], aes(x=numericFactorOrder(TransferTime), y = ratio, group = Set, fill = Substrate, colour = Substrate, shape=Set)) +
  geom_point(size = 2, position = position_dodge(0.3)) +
  geom_errorbar(aes(ymin = ratio-se, ymax = ratio+se), width = 0.08, position = position_dodge(0.3)) +
  scale_shape_manual(values = c(21, 22, 23, 24, 25)) +
  scale_color_manual(values = c('#a6cee3','#1f78b4'), guide = FALSE)  +
  scale_fill_manual(values = c('#a6cee3','#1f78b4'), guide = FALSE)  +
  #ylim(c(0,1)) +
  labs(shape = "Researcher",
       x = "Transfer Time (s)",
       y = "Transfer Ratio") +
  facet_grid(Substrate ~ .) +
  mytheme +
  theme_pubr()

#trp1 

t60.dat =  all.ratio %>% filter(TransferTime == 60 & Substrate %in% c("Wool", "Nylon"))

trp2 = ggplot(t60.dat, aes(x=numericFactorOrder(Mass), y = ratio, group = Set, shape=Set, fill = Substrate, colour=Substrate)) +
  geom_point(size = 2, position = position_dodge(0.3)) +
  geom_errorbar(aes(ymin = ratio-se, ymax = ratio+se), width = 0.08, position = position_dodge(0.3)) +
  scale_shape_manual(values = c(21, 22, 23, 24, 25)) +
  scale_color_manual(values = c('#a6cee3','#1f78b4'), guide = FALSE)  +
  scale_fill_manual(values = c('#a6cee3','#1f78b4'), guide = FALSE)  +
  labs(shape = "Researcher",
       x = "Mass (g)",
       y = "Transfer Ratio") +
  facet_grid(Substrate ~ .) +
  #mytheme +
  theme_pubr()

trp12 = ggarrange(trp1, trp2, labels = c('A', 'B'), ncol = 2,
          common.legend = TRUE, legend = 'top')
trp12

#ggsave('transratio.png', trp12,  height = 4.5, units = 'in')
```

Figure S18: Transfer ratios for all datasets. (A) for mass = 1000 g and all times, (B) for time = 60 s and all masses.

```{r plot60w, eval = FALSE,fig.height= 3.9}

wool60.dat = all.ratio %>% filter(TransferTime == 60 & Substrate == "Wool")
#wool60.dat

ggplot(wool60.dat, aes(x=numericFactorOrder(Mass), y = ratio, group = Set, colour = Set, shape=Set)) +
  geom_point(position = position_dodge(0.3)) +
  geom_errorbar(aes(ymin = ratio-se, ymax = ratio+se), width = 0.08, position = position_dodge(0.3)) +
  scale_shape_manual(values = c(0, 1, 2, 5, 6)) +
  ylim(c(0,1)) +
  labs(title = "Transfer Ratio for Cotton to Wool",
       subtitle = "For 60s Transfer Time",
       x = "Mass (g)",
       y = "Transfer Ratio") +
  mytheme

#ggsave("fig3.png")

```
```{r plot60n, eval = FALSE}

nylon60.dat = all.ratio %>% filter(TransferTime == 60 & Substrate == "Nylon")
#nylon60.dat

ggplot(nylon60.dat, aes(x=numericFactorOrder(Mass), y = ratio, group = Set, colour = Set, shape=Set)) +
  geom_point(position = position_dodge(0.3)) +
  geom_errorbar(aes(ymin = ratio-se, ymax = ratio+se), width = 0.08, position = position_dodge(0.3)) +
  scale_shape_manual(values = c(0, 1, 2, 5, 6)) +
  labs(title = "Transfer Ratio for Cotton to Nylon",
       subtitle = "For 60s Transfer Time",
       x = "Mass (g)",
       y = "Transfer Ratio") +
  mytheme

#ggsave("fig4.png")

```


```{r,fig.height= 3.9}

# plot boxplots
ggplot(all.ratio %>% 
  filter(Substrate %in% c('Wool', 'Nylon')), aes(x=Set, y=ratio, fill = Substrate)) +
  geom_boxplot() +
  scale_fill_manual(values = c('#a6cee3','#1f78b4')) +
  labs(x = "Researcher",
       y = "Transfer Ratio") +
  mytheme +
  theme_pubr()

#ggsave("transratio_boxplot.png", height = 4, width=6, units = 'in')
#kruskal.test(Ratio ~ Substrate, data = all.ratio.tmp)
```

Figure S19: Transfer ratios for Wool and Nylon, variation between dataset (researcher).

```{r,fig.height= 4}

ggplot(all.dat, aes(x=P3)) + geom_density() +
  labs(title = "Distribution of Particle Counts",
       substitle = "Photo 3",
       x = "") +
  facet_grid(Set ~ ., scales = "free_y")

```

Figure S20: Particle count distribution for Photo 3 (donor after UV deposition) between datasets.

```{r plotallratio, fig.height=3.9,fig.width= 8}

# for better labelling add a new mass column with a 'g'
all.ratio$Massg = paste0(all.ratio$Mass, "g")

ggplot(all.ratio, aes(x = numericFactorOrder(TransferTime), y = ratio, group = Set, colour = Set)) + 
  geom_point(size=1, position = position_dodge(0.3)) +
  geom_errorbar(aes(ymin = ratio-se, ymax = ratio+se), width = 0.08, position = position_dodge(0.3)) +
#  geom_point(colour = "white", size = 0.7, position = position_dodge(0.2)) +
#  scale_x_discrete(labels = c("Wool", "Nylon")) +
  labs(title = "Analysis of Transfer Ratio",
       x = "Transfer Time (s)",
       y = "Transfer Ratio") +
  facet_grid(fct_reorder(Massg, Mass) ~ Substrate) +
  mytheme +
  theme(strip.background = element_rect(fill = "lightblue", colour = "black"))


```

Figure S21: Summary of all transfer ratios.

\newpage
# Transfer Efficiencies
```{r ploteff,fig.height= 3.9}
# some replicate efficiencies are infinite 
# remove them now
all.final = all.final[is.finite(all.final$Efficiency),]

# first calc the efficiency data
all.eff = all.final %>% 
  group_by(Set, Mass, TransferTime, Substrate) %>%
  summarise(N = n(), efficiency = mean(Efficiency), se = stdErr(Efficiency))

eff1000.dat = all.eff %>% filter(Mass == 1000 & TransferTime != "10")
#nylon.eff.dat

tep1 = ggplot(eff1000.dat[eff1000.dat$Substrate %in% c('Wool', 'Nylon'),], aes(x=numericFactorOrder(TransferTime), y = efficiency, group = Set, shape=Set, colour = Substrate, fill = Substrate)) +
  geom_point(size = 2, position = position_dodge(0.3)) +
  geom_errorbar(aes(ymin = efficiency-se, ymax = efficiency+se), width = 0.08, position = position_dodge(0.3)) +
  scale_shape_manual(values = c(21, 22, 23, 24, 25)) +
  scale_color_manual(values = c('#a6cee3','#1f78b4'), guide = FALSE)  +
  scale_fill_manual(values = c('#a6cee3','#1f78b4'), guide = FALSE)  +
  labs(shape = "Researcher",
       x = "Transfer Time (s)",
       y = "Transfer Efficiency") +
  facet_grid(Substrate ~ .) +
  theme_pubr()

#tep1
```


```{r ploteff60,fig.height= 3.9}
eff60.dat = all.eff %>% filter(TransferTime == 60 & Substrate != "Denim")

tep2 = ggplot(eff60.dat, aes(x=numericFactorOrder(Mass), y = efficiency, group = Set, shape=Set, colour = Substrate, fill = Substrate)) +
  geom_point(position = position_dodge(0.3)) +
  geom_errorbar(aes(ymin = efficiency-se, ymax = efficiency+se), width = 0.08, position = position_dodge(0.3)) +
  scale_shape_manual(values = c(21, 22, 23, 24, 25)) +
  scale_color_manual(values = c('#a6cee3','#1f78b4'), guide = FALSE)  +
  scale_fill_manual(values = c('#a6cee3','#1f78b4'), guide = FALSE)  +
  labs(Shape = "Researcher",
       x = "Mass (g)",
       y = "Transfer Efficiency") +
  facet_grid(Substrate ~ .) +
  theme_pubr()

tep12 = ggarrange(tep1, tep2, labels = c('A', 'B'), ncol = 2,
                  common.legend = TRUE)
tep12
#ggsave("transeff.png", tep12, height = 4, width = 6, units = 'in')


```

Figure S22: Transfer efficiencies for all datasets. (A) for mass = 1000 g and all times, (B) for time = 60 s and all masses.

```{r plotalleff, fig.height=3.9,fig.width= 8}

all.eff$Massg = paste0(all.eff$Mass, "g")

ggplot(all.eff, aes(x = numericFactorOrder(TransferTime), y = efficiency, group = Set, colour = Set)) + 
  geom_point(size=1, position = position_dodge(0.3)) +
  geom_errorbar(aes(ymin = efficiency-se, ymax = efficiency+se), width = 0.08, position = position_dodge(0.3)) +
#  geom_point(colour = "white", size = 0.7, position = position_dodge(0.2)) +
#  scale_x_discrete(labels = c("Wool", "Nylon")) +
  labs(title = "Analysis of Transfer Efficiency",
       x = "Transfer Time (s)",
       y = "Transfer Efficiency") +
  facet_grid(fct_reorder(Massg, Mass) ~ Substrate) +
  mytheme +
  theme(strip.background = element_rect(fill = "lightblue", colour = "black"))


```

Figure S23: Summary of all transfer efficiencies.